---
title: 'Reproducible Research: Peer Assessment 1'
author: "Aashish Shetty"
output:
  html_document:
    keep_md: true
    pdf_document: default
---

### Report generated on `r format(Sys.Date(), "%B %d, %Y")`

```{r global_options, include=FALSE}
knitr::opts_chunk$set( echo=TRUE, warning=FALSE, message=FALSE, fig.keep="all")
#Set scipen to 100 to prevent values being print as a power of 10
options("scipen"=100, "digits"=4)
```

## Loading and preprocessing the data
```{r load_preprocess}
    #Load Libraries needed
    library(data.table)
    library(dplyr)
    library(lattice)

    #Load the data (i.e. read.csv())
    csv<-read.csv("activity.csv",as.is=F )

    #Transform  data to data table to help with analysis
    data_table<-data.table(csv)
```

***  


## What is mean total number of steps taken per day?
```{r steps_summary}
    by_date<-group_by(data_table,date)
    steps_summary<-summarise(by_date,step_count=sum(steps,na.rm=T))
    histogram(~step_count,
        data=steps_summary,
        type="count",
        ylim=c(0,12),
        breaks=40,
        xlab="No. of Steps per day",
        main="Histogram of no. of steps per day")

#Mean of number of steps taken each day is  `r mean(steps_summary$step_count)`  
#Median of number of steps taken each day is `r median(steps_summary$step_count)`   
```

####Mean of number of steps taken each day is  `r mean(steps_summary$step_count)`  
####Median of number of steps taken each day is `r median(steps_summary$step_count)`   
  
***  
  
## What is the average daily activity pattern?

```{r avg_daily_pattern}

by_interval<-group_by(data_table,interval)
interval_summary<-summarise(by_interval,interval_avg=mean(steps,na.rm=T))

xyplot( interval_avg ~ interval,
    data = interval_summary,
    type = "l",
    xlab= "5- minute Interval", 
    ylab= "Average number of steps take in an interval across all days",
    main="Time series plot:average steps of the 5-minute interval, across all days",
    col.line = c(rep("blue",3), "red"))

####On average across all the days in the dataset, the 5-minute interval `r interval_summary[which.max(interval_avg),]$interval ` contains the maximum number of steps

####Total number of missing values in the dataset `r length(which(is.na(data_table$steps)))`

```


####On average across all the days in the dataset, the 5-minute interval `r interval_summary[which.max(interval_avg),]$interval ` contains the maximum number of steps
    
***  

## Imputing missing values

####Total number of missing values in the dataset `r length(which(is.na(data_table$steps)))`

###Strategy for Imputing missing values
####Replace missing values with the with mean for that 5-minute interval. This has been saved in interval_summary in the previous step

```{r imputing_missing_values, cache=F}

new_data_table<-copy(data_table)

for(i in which(is.na(new_data_table$steps)))
{
    new_data_table[i]$steps = 
        as.integer(interval_summary[interval==new_data_table[i]$interval]$interval_avg)
}

by_date2<-group_by(new_data_table,date)
steps_summary2<-summarise(by_date2,step_count=sum(steps,na.rm=T))

histogram(~step_count,
    data=steps_summary2,
    type="count",
    ylim=c(0,12),
    breaks=40,
    xlab="No. of Steps per day",
    main="Histogram of no. of steps per day after imputing missing values")


####Before imputing data Mean of no. of steps was `r mean(steps_summary$step_count)`  , after imputing missing data it has increased to `r mean(steps_summary2$step_count)`   
####Before imputing data Median of no. of steps is `r median(steps_summary$step_count)`, after 
```


####Before imputing data Mean of no. of steps was `r mean(steps_summary$step_count)`  , after imputing missing data it has increased to `r mean(steps_summary2$step_count)`   
####Before imputing data Median of no. of steps is `r median(steps_summary$step_count)`, after imputing missing data it has increased to `r median(steps_summary2$step_count)`

####Hence after imputing missing data, the median and mean increased slightly 
***  

## Are there differences in activity patterns between weekdays and weekends?

```{r weekend_activity}

isWeekday<-function(x){
    if(grepl("Sunday|Saturday",x))
        return("Weekend")
    else
        return("Weekday")
}

#Add Weekday/Weekend - day of the week column as day
data_table<-data_table[,day:=sapply(weekdays(as.Date(date , format = "%Y-%m-%d")),isWeekday)]

day_interval<-group_by(data_table,day,interval)
week_summary<-summarise(day_interval,interval_avg=mean(steps,na.rm=T))

xyplot( interval_avg ~ interval| day,
    data = week_summary,
    type = "l",
    lty = c(1, 2, 2, 1),
    lwd = c(1, 1, 1, 3),
    layout=c(1,2),
    xlab= "5- minute Interval", 
    ylab= "Average number of steps take in an interval across all days",
    col.line = c(rep("blue",3), "red"),
    main="Comparison between Weekday and weekend activity")

```


