---
title: 'Reproducible Research: Peer Assessment 1'
author: "Aashish Shetty"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
<!--
#`{r} knitr::opts_chunk$set(message=F,echo = FALSE, results = "hide", cache=FALSE)  `
date: "Friday, May 15, 2015"

#```{r, echo=FALSE} knitr::opts_chunk$set(results="hide", fig.show="hide") ```
-->

### Report generated on `r format(Sys.Date(), "%B %d, %Y")`

```{r global_options, include=FALSE}
knitr::opts_chunk$set( echo=FALSE, warning=FALSE, message=FALSE)
#Set scipen to 100 to prevent values being print as a power of 10
options("scipen"=100, "digits"=4)
```

```{r Load_libraries_needed}
#Load Libraries needed
library(data.table)
library(dplyr)
library(ggplot2)
library(lattice)
```

## Loading and preprocessing the data
```{r load_preprocess}
    #Load the data (i.e. read.csv())
    #csv<-read.csv("activity.csv",colClasses=c("numeric","Date","numeric",NA) )
    csv<-read.csv("activity.csv",as.is=F )

    #Transform  data to data table to help with analysis
    data_table<-data.table(csv)
    summary(data_table)
```



## What is mean total number of steps taken per day?
```{r steps_summary}
    by_date<-group_by(data_table,date)
    steps_summary<-summarise(by_date,step_count=sum(steps,na.rm=T))
    hist(steps_summary$step_count,
         ylim=c(0,10),
         breaks=40,
         xlab="No. of Steps per day",
         main="Histogram of no. of steps per day")
steps_summary
str(steps_summary)

```

###Mean of no. of steps is `r mean(steps_summary$step_count)`  
###Median of no. of steps is `r median(steps_summary$step_count)`   
  
***  
  
## What is the average daily activity pattern?

```{r avg_daily_pattern}
    by_interval<-group_by(data_table,interval)
    interval_summary<-summarise(by_interval,interval_avg=mean(steps,na.rm=T))
    plot(interval_summary$interval_avg, 
         type="l", 
         xlab= "Interval", 
         ylab= "Average number of steps take in an interval across all days",
         main="Time series plot:average steps of the 5-minute interval, across all days",
         )
    
```


###On average across all the days in the dataset, interval `r interval_summary[which.max(interval_avg),]$interval ` contains the maximum number of steps
    
***  
<!--
```{r}


    by_date<-group_by(data_table,date)
    steps_summary<-summarise(by_date,step_count=sum(steps,na.rm=T))
    hist(steps_summary$step_count,
         ylim=c(0,15),
         breaks=40,
         xlab="No. of Steps per day",
         main="Histogram of no. of steps per day")
```
-->

## Imputing missing values
```{r imputing_missing_values, cache=TRUE}

na_rows<-filter(data_table,is.na(steps) )
nrow(na_rows)
new_data_table<-copy(data_table)

for(i in 1:nrow(new_data_table))
    {
        if(is.na(new_data_table[i]$steps))
        {
            new_data_table[i]$steps = 
                as.integer(interval_summary[interval==new_data_table[i]$interval]$interval_avg)
        }
    }
by_date2<-group_by(new_data_table,date)
steps_summary2<-summarise(by_date2,step_count=sum(steps,na.rm=T))
hist(steps_summary2$step_count,
     ylim=c(0,15),
     breaks=40,
     xlab="No. of Steps per day",
     main="Histogram of no. of steps per day")

```


###Mean of no. of steps is `r mean(steps_summary$step_count)`  
###Median of no. of steps is `r median(steps_summary$step_count)`   
### Means of no. of steps is `r mean(steps_summary2$step_count)`  
###Median of no. of steps is `r median(steps_summary2$step_count)`   

***  

## Are there differences in activity patterns between weekdays and weekends?

```{r weekend_activity}

isWeekday<-function(x){
    if(grepl("Sunday|Saturday",x))
        return("Weekend")
    else
        return("Weekday")
}

#Add day of the week column as day
data_table[,day:=sapply(weekdays(as.Date(date , format = "%Y-%m-%d")),isWeekday)]

day_interval<-group_by(data_table,day,interval)
week_summary<-summarise(day_interval,interval_avg=mean(steps,na.rm=T))
xyplot( interval_avg ~ interval| day,
    data = week_summary,
    type = "l",
    lty = c(1, 2, 2, 1),
    lwd = c(1, 1, 1, 3),
    layout=c(1,2),
    col.line = c(rep("blue",3), "red"))

#     we_interval<-group_by(weekend,interval)
#     weekend_summary<-summarise(we_interval,interval_avg=mean(steps,na.rm=T))
#     plot(weekend_summary$interval_avg, 
#          type="l", 
#          xlab= "Interval", 
#          ylab="Average number of steps take in an interval across all weeknds",
#          main="Time series plot:average steps of the 5-minute interval, across all weekends",
#          )
#     
#     wd_interval<-group_by(weekday,interval)
#     weekday<-summarise(wd_interval,interval_avg=mean(steps,na.rm=T))
#     plot(weekday$interval_avg, 
#          type="l", 
#          xlab= "Interval", 
#          ylab= "Average number of steps take in an interval across all days",
#          main="Time series plot:average steps of the 5-minute interval, across all days",
#          )
#     

```


